Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fix issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/41 and increase font also in Fig. 5
Small change to test PR.

The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral calculated from different models are
shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger-POPC-07 model, showing a maximum in 
the gauche$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the anti-conformation (180$^o$).
The distributions in all the other models have the same general features, the main difference being that the
fraction of configurations in the gauche$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
equally large to the gauche$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
this is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.pdf}
  \caption{\label{dihDISTS}
    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral from different models (POPC bilayer in full hydration).
      } 
\end{figure}

@@

%Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
%the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
%Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
%larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
%the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
%model suggesting that it has some other inaccuracies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.pdf}
  \caption{\label{dihDISTS2}
    Dihedral angle distributions for N-$\beta$-$\alpha$-O dihedral from different models (POPC bilayer in full hydration).
  } 
\end{figure}

@@

The qualitative agreement with experiments in all simulation models for the $\alpha$ and $\beta$ order parameters  
as a function of hydration indicates that the structural response of the choline headgroup to dehydration is somewhat realistic
despite the unrealistic structures at full hydration. 
The most likely explanation is that the choline group
orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
Indeed, the P--N (phoshate phosphorus to choline nitrogen) vector angle with membrane normal shows an increase for
all models as a function of dehydration in Fig.~\ref{PNangle}.
However, the amount of increase depends on the model. Especially the DLPC simulations with Berger model
predict significantly stronger P--N vector tilt compared with the other models. The Berger model
has also generally larger P--N vector angles and its choline order parameters are more off from
experiments than other models. Thus the relatively modest tilting with dehydration
predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.

>>>>

The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral calculated from different models are
shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger-POPC-07 model, showing a maximum in 
the gauche$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the anti-conformation (180$^o$).
The distributions in all the other models have the same general features, the main difference being that the
fraction of configurations in the gauche$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
equally large to the gauche$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
this is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.eps}
  \caption{\label{dihDISTS}
    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral from different models (POPC bilayer in full hydration).
      } 
\end{figure}

@@

%Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
%the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
%Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
%larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
%the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
%model suggesting that it has some other inaccuracies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.eps}
  \caption{\label{dihDISTS2}
    Dihedral angle distributions for N-$\beta$-$\alpha$-O dihedral from different models (POPC bilayer in full hydration).
  } 
\end{figure}

@@

The qualitative agreement with experiments in all simulation models for the $\alpha$ and $\beta$ order parameters  
as a function of hydration indicates that the structural response of the choline headgroup to dehydration is somewhat realistic
despite the unrealistic structures at full hydration. 
The most likely explanation is that the choline group
orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
Indeed, the P--N (phoshate phosphorus to choline nitrogen) vector angle with membrane normal shows an increase for
all models as a function of dehydration (Fig.~\ref{PNangle}). However,
the amount of increase depends on the model. Especially the DLPC simulations with Berger model
predict significantly stronger P--N vector tilt compared with the other models. The Berger model
has also generally larger P--N vector angles and its choline order parameters are more off from
experiments than other models. Thus the relatively modest tilting with dehydration
predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.
